Temporal dataflows and their application to generative music

### 2017-03-22 article-icmc2017.tex
save

\usepackage{icmc2017template}
\usepackage{times}
\usepackage{ifpdf}
\usepackage{soul}
\usepackage{microtype}
\usepackage{polyglossia}
%\usepackage{geometry}
\usepackage{tikz}
\usepackage{listings}

\def\papertitle{Paper Template for ICMC 2017}
\def\firstauthor{First Author}
\def\secondauthor{Second Author}
\def\thirdauthor{Third Author}
\def\fourthauthor{Fourth Author}
\def\fifthauthor{Fifth Author}
\def\sixthauthor{Sixth Author}

@@

\title{Temporal dataflow graphs and their application to generative music}

@@

    \capstartfalse
    \maketitle
    \capstarttrue

@@

    The main problem that arises is that composers may specify relationships between processes, but such processes may not always be active. 
    We start by making the case for an environment associated with the dataflow graph that contains the values of the inputs and outputs.
    Then, since a graph may not always be fully active, we show that it is also meaningful to have implicit connections between nodes of the graph, that will use the environment.
    This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
    Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.
    \subsection{Related works}
    Citer Kyma
    Céu
    Perl

@@

	\section{Graph definitions}
    We rely on two families of graphs for the execution of the software: 
    \begin{itemize}
        \item A dataflow graph between processes: processes have input and output ports that can be connected together to enforce data dependencies.
        \item A temporal graph, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score.
    \end{itemize}
    There may be more than a single temporal graph: they can be arranged hierarchically.

    The execution is driven by a scheduler.

    At each tick, the state of each temporal graph is updated with the following informations: 
    \begin{itemize}
        \item Which processes are currently running.
        \item For how long these processes have been running.
    \end{itemize}

@@

    Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound. (Do like LV2: pass empty buffers). Pb: multi-rate ... voir méthode pour passer de multi-rate à single rate dans PLOP2006 et Signal Flow Graphs and Data Flow Graphs (3.2, 3.3), dans Handbook of Signal Processing (First Edition)

@@

    Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example. 
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=0.5]
        \draw[line width=2pt] (0, 0) -- (0, 2) -- (5, 2) -- (5, 0) -- (0, 0);
        \draw[line width=2pt] (0 + 3, 0 - 2.5) -- (0 + 3, 2 - 2.5) -- (5 + 3, 2 - 2.5) -- (5 + 3, 0 - 2.5) -- (0 + 3, 0 - 2.5);

@@

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=0.5]
        \draw[line width=2pt] (0 + 3, 0) -- (0 + 3, 2) -- (5 + 3, 2) -- (5 + 3, 0) -- (0 + 3, 0);
        \draw[line width=2pt] (0, 0 - 2.5) -- (0, 2 - 2.5) -- (5, 2 - 2.5) -- (5, 0 - 2.5) -- (0, 0 - 2.5);

@@

    We extend this with a permanent environment, that contains a mapping of key-value pairs. 
    Keys are adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed through the network, graphical or physical controls.

>>>>

\usepackage{icmc2017template}
\usepackage{amssymb}
\usepackage{times}
\usepackage{ifpdf}
\usepackage{soul}
\usepackage{microtype}
\usepackage{polyglossia}
%\usepackage{geometry}
\usepackage{tikz}
\usepackage{listings}

\def\papertitle{Paper Template for ICMC 2017}
\def\firstauthor{First Author}
\def\secondauthor{Second Author}
\def\thirdauthor{Third Author}

\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}

@@

\title{Temporal dataflows and their application to generative music}

@@

    \capstartfalse
    \maketitle
    \capstarttrue
    \begin{abstract}
        contenu...
    \end{abstract}

@@

    The main problem that arises is that composers may specify relationships between processes, but such processes may not always be active. 
    We start by making the case for an environment associated with the dataflow graph that contains the values of the inputs and outputs.
    Then, since a graph may not always be fully active, we show that it is also meaningful to have implicit connections between nodes of the graph, that will use the environment.
    This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
    Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.

    \subsection{Related works}
    Citer Kyma
    Céu
    Perl
    Boolean dataflows (http://pop-art.inrialpes.fr/~fradet/PDFs/RR8333.pdf)

@@

    \section{Definitions}
    We call process an entity that consists in a state, an associated function $f$, any number of inputs and output ports, and an activation status.
    $f$ has a single argument $t \in \mathbb{R^+}$.
    The execution of $f$ may read any number of tokens from the inputs and write any number of tokens to the outputs.

    We rely on two families of graphs for defining the execution of a program:

    \begin{itemize}
        \item Synchronous dataflows between processes: input and output ports can be connected together to enforce data dependencies.
        \item Temporal graphs, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score. 
        For coherency, we call these graphs timeflows.
    \end{itemize}

    A program consists in the sequential and parallel execution of multiple processes, specified by a single dataflow, and at least one timeflow.

    There may be more than a single timeflow: they are defined as processes themselves, which provides them with time information.
    They have no input or output.

    Each time constraint has an associated set of references to processes.
    For each process in the dataflow, there is exactly one reference to this process in the set of all timeflows.

    The execution of a program is synchronous and driven by a scheduler.
    At each tick, the state of each temporal graph is updated with the following informations:

    \begin{itemize}
        \item Which processes are currently activated.
        \item For how long these processes have been activated.
    \end{itemize}

@@

    Only activated processes may be called in a given tick.

    % In a given tick, every active process will be executed. ?
    %Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound. (Do like LV2: pass empty buffers). Pb: multi-rate ... voir méthode pour passer de multi-rate à single rate dans PLOP2006 et Signal Flow Graphs and Data Flow Graphs (3.2, 3.3), dans Handbook of Signal Processing (First Edition)

@@

    Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example. 
    \begin{figure}
        \centering
        \begin{tikzpicture}[scale=0.5]
        \draw[line width=2pt] (0, 0) rectangle (5, 2);
        \draw[line width=2pt] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5);

@@

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=0.5]

        \draw[line width=2pt] (0 + 3, 0) rectangle (5 + 3, 2);
        \draw[line width=2pt] (0, 0 - 2.5) rectangle (5, 2 - 2.5);

@@

    We extend this with a permanent external environment, that contains a mapping of key-value pairs. 
    Keys are specified with OSC-like\cite{osc} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed asynchronously through the network, graphical widgets or physical controls.
